Chapter3_ideas

### 2019-12-09 results/Chapter3_ideas.Rmd
rbar for rand and norm foci pos

#read files
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
```

@@

whole.cell.data <- add_IFD(whole.cell.data)
whole.cell.data <-  add_IFD.2(whole.cell.data)

Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
#IFD.2 does mulitple IFDs (just absolute)
Curated_BivData <- add_IFD.2(Curated_BivData)

@@

1. use the same randomized dataframe
2. use the summary stat

@@

pos.pool_1CO <- male.whole.cell[male.whole.cell$hand.foci.count == 1,]
pos.pool_2CO <- male.whole.cell[male.whole.cell$hand.foci.count == 2,]
pos.pool_3CO <- male.whole.cell[male.whole.cell$hand.foci.count == 3,]

#now make the randomized data set for CO pos
#do whole row -- so I can cen n

random_pos <- data.frame(male.whole.cell$fileName, male.whole.cell$SC.length, male.whole.cell$hand.foci.count, male.whole.cell$Obj.ID  )

#INITIALIZE EMPTY list of DFs
#fill in with full DF

LIST.DFs <- setNames( replicate( length(male.biv.count_table$fileName), data.frame( obj.ID = c())), male.biv.count_table$fileName)

ids <-    male.whole.cell$Obj.ID[which(male.whole.cell$fileName == '14oct16_30jun16_CAST_m3_sp1_3.2')]

LIST.DFs[1]$obj.ID[1:19] <- ids[1:19]

#INITALIZE empty LIST (make DF in the loop)

LIST <- list()

#Loop for rand sampling  
for( male.cell in male.biv.count_table$fileName){

  #for each fileName, make a new DF
  DF.in.loop <- data.frame(matrix(ncol = 1, nrow = 19))
  #making the Within loop DF
  DF.in.loop$X1 <- male.whole.cell$fileName[male.whole.cell$fileName ==male.cell]
  DF.in.loop$obj.ID <- male.whole.cell$Obj.ID[male.whole.cell$fileName ==male.cell]
  DF.in.loop$SC.length <- male.whole.cell$SC.length[male.whole.cell$fileName ==male.cell]
  DF.in.loop$hand.foci.count <- male.whole.cell$hand.foci.count[male.whole.cell$fileName ==male.cell]

  #LIST.DFs[[male.cell]]$obj.ID[1:19] <- male.whole.cell$Obj.ID[male.whole.cell$fileName == male.cell]
    #print(male.cell)
    n0CO = male.biv.count_table$nCO0[male.biv.count_table$fileName == male.cell] #this reference needs to be from male.biv.count_table
    n1CO = male.biv.count_table$nCO1[male.biv.count_table$fileName == male.cell]
    n2CO = male.biv.count_table$nCO2[male.biv.count_table$fileName == male.cell]
    n3CO = male.biv.count_table$nCO3[male.biv.count_table$fileName == male.cell]

   #rand samples for foci.pos
    #JUST DO THIS FOR 1CO bivs
    DF.in.loop$rand_F1
    sample.n1CO <- sample(pos.pool_1CO$Foci1.PER, n1CO, replace = FALSE )  #sample from the cell.table

    #sample a random 2CO chrm (obj.ID),    #doesn't make sense to sample a random position from 2CO pool, need to sample the linked positions, or get the 
    #chrm observation
   sample.n2CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n2CO, replace = FALSE )
   sample.n3CO.obj.id <- sample(pos.pool_2CO$Obj.ID, n3CO, replace = FALSE )

   #MAP normalized positions onto SC lengths/chrms of the fileName cell obs
   #map the 1CO rand Foci position onto 1CO SCLength

   #for i in n1CO
#   which(random_pos$fileName == male.cell)
   random_pos$Rand.F1[which(random_pos$fileName == male.cell)][i] <- sample(pos.pool_1CO$Foci1.PER, 1, replace = FALSE)

   #list of the chrm lengths for all the 1COs for the current cell obs
   #SC.Length_1COs <-

@@

#need to define number of chrm classes within each nuc
n1CO = 9

BIG.DF.LIST_COpos_no.rank <- replicate(10,
                  list(bind_rows(list(

  pos.pool_1CO[ sample( which( pos.pool_1CO$hand.foci.count == 1), n1CO, replace = FALSE ),]

) ) ) )

BIG.DF.LIST_no.rank <- replicate(18,
                  list(bind_rows(list( 
          male.whole.cell[ sample( which( male.whole.cell$rank < 20), 19, replace = FALSE ),]
) ) ) )

@@

>>>>

#read files
whole.cell.data <- read.csv("~./MLH1repo/whole_cell_measures.csv")
#load(file="~./MLH1repo/data/CleanBivData.RData")#added batch17
```

@@

whole.cell.data <- add_IFD(whole.cell.data)
#whole.cell.data <-  add_IFD.2(whole.cell.data)

#Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength
#IFD.2 does mulitple IFDs (just absolute)
#Curated_BivData <- add_IFD.2(Curated_BivData)

@@

1. use the same randomized dataframe
2. use the summary stat
3. draw rand nrm foci position ... then map that onto the raw SC length

@@

male.whole.cell <- male.whole.cell[!(is.na(male.whole.cell$fileName) | male.whole.cell$fileName==""), ]

#make sure male.whole.cell has nrm oci positions
pos.pool_1CO <- male.whole.cell[male.whole.cell$hand.foci.count == 1,]
pos.pool_2CO <- male.whole.cell[male.whole.cell$hand.foci.count == 2,]
pos.pool_3CO <- male.whole.cell[male.whole.cell$hand.foci.count == 3,]

pos.pool_1CO <- pos.pool_1CO[!(is.na(pos.pool_1CO$fileName) | pos.pool_1CO$fileName==""), ]
pos.pool_2CO <- pos.pool_2CO[!(is.na(pos.pool_2CO$fileName) | pos.pool_2CO$fileName==""), ]
pos.pool_3CO <- pos.pool_3CO[!(is.na(pos.pool_3CO$fileName) | pos.pool_3CO$fileName==""), ]
#now make the randomized data set for CO pos
#do whole row -- so I can cen n

#make a copy of main DF
male.whole.cell.mini <- male.whole.cell#[190:455, ]

for(oar in 1:length(male.whole.cell.mini$fileName)){
 # print(oar)
   if(male.whole.cell.mini$hand.foci.count[oar] == 1){
    print("oner")
     #random sample -- there could be replicates in the random sample
     rand.obj.id.1CO <- sample(pos.pool_1CO$Obj.ID, 1)

     male.whole.cell.mini$rand.ID[oar] <- rand.obj.id.1CO
     male.whole.cell.mini$rand.Foci1[oar] <- as.numeric(pos.pool_1CO$Foci1.PER[which(pos.pool_1CO$Obj.ID == rand.obj.id.1CO)])
     #  sample(pos.pool_1CO$F1, 1, replace = FALSE)
     male.whole.cell.mini$rand.Foci2[oar] <- ""
     male.whole.cell.mini$rand.Foci3[oar] <- ""
   }

  if(male.whole.cell.mini$hand.foci.count[oar] == 2){
   # print("twoer")

    #random sampl of obj.id
    rand.obj.id <- sample(pos.pool_2CO$Obj.ID, 1)

    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id
    #asign the F1s and F2s from the fit obj.id

    male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_2CO$Foci1.PER[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_2CO$Foci2.PER[which(pos.pool_2CO$Obj.ID == rand.obj.id)])
    male.whole.cell.mini$rand.Foci3[oar] <- ""
 }

  if(male.whole.cell.mini$hand.foci.count[oar] == 3){
   # print("threer")
    rand.obj.id_3co <- sample(pos.pool_3CO$Obj.ID, 1)

    male.whole.cell.mini$rand.ID[oar] <- rand.obj.id_3co

        male.whole.cell.mini$rand.Foci1[oar] <-as.numeric(pos.pool_3CO$Foci1.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])
  #  male.whole.cell.mini$rand.Foci2[oar] <- "blah"
    male.whole.cell.mini$rand.Foci2[oar] <-as.numeric(pos.pool_3CO$Foci2.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])

    male.whole.cell.mini$rand.Foci3[oar] <-as.numeric(pos.pool_3CO$Foci3.PER[which(pos.pool_3CO$Obj.ID == rand.obj.id_3co)])

@@

@@

### merge daata

```{r merging.data,echo=FALSE}

#melt the nrmfoci.pos, for real and randomized data
REAL.melt.whole.cell.mini <- melt(male.whole.cell.mini,
              c("Obj.ID","SC.length", "hand.foci.count","Foci1.PER","Foci2.PER","Foci3.PER"),

              id=c("Obj.ID","SC.length", "hand.foci.count")
                                  )
REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[!(is.na(REAL.melt.whole.cell.mini$value) | REAL.melt.whole.cell.mini$value==""), ]

REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "hand.foci.count"), ]
REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "Obj.ID"), ]
REAL.melt.whole.cell.mini <- REAL.melt.whole.cell.mini[ (REAL.melt.whole.cell.mini$variable != "SC.length"), ]

#make random melted DF
RAND.melt.whole.cell.mini <- melt(male.whole.cell.mini, c("rand.ID","SC.length","hand.foci.count","rand.Foci1","rand.Foci2","rand.Foci3"),
                                  id=c("rand.ID", "SC.length","hand.foci.count")
)
RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[!(is.na(RAND.melt.whole.cell.mini$value) | RAND.melt.whole.cell.mini$value==""), ]

RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "hand.foci.count"), ]
RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "rand.ID"), ]
RAND.melt.whole.cell.mini <- RAND.melt.whole.cell.mini[ (RAND.melt.whole.cell.mini$variable != "SC.length"), ]

```{r merge.real.rand, echo=FALSE}

REAL.melt.whole.cell.mini$type <- "real"
RAND.melt.whole.cell.mini$type <- "sim"

colnames(RAND.melt.whole.cell.mini) <- c("Obj.ID","SC.length", "hand.foci.count","variable","value", "type" )

Position.DF <- rbind(REAL.melt.whole.cell.mini,RAND.melt.whole.cell.mini)

Position.DF$value <- as.numeric(Position.DF$value )
Position.DF$SC.length <- as.numeric(Position.DF$SC.length )

### plot relative foci positions

```{r Pos.plot, echo=FALSE}

#density or histogram plot facet hand foci count, plot value, fill type

pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)

pos.plot <- ggplot(Position.DF[Position.DF$hand.foci.count == 1,], aes(x=as.numeric(value), fill=type))+geom_histogram()

pos.plot.all <- ggplot(Position.DF, aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count,ncol = 1)+ggtitle("Simulated and normalized positions, male mus")

pos.plot.all

pos.plot.one <- ggplot(Position.DF[Position.DF$hand.foci.count == 2,], aes(x=as.numeric(value), fill=type))+geom_density(alpha=.6)+facet_wrap(~hand.foci.count)+ggtitle("Simulated and normalized positions")

HS.density.plot <- ggplot(Hs.CO_cVs.DF, aes(x = COs, fill=type) )+geom_density(alpha=.6) + ggtitle("CO per cell, Human male")+xlim(c(0,100))

The above distributions of relative foci position compare the real data pink to randomized data set of relative positions where for each bivalent observations the normalized position was replaced with positions from a randomly draw bivalent with the same CO number. For multi co bivalents relative foci positions were kept together. For example, a random 2CO bivalent choosen and the normalize foci1 and foci2 for that bivalent were used.

In terms of normalized foci positions, the empiracal distribution isn't very different from a randomized distribution.

Not exactly sure if this is the right randomization procedure, the same 'random' bivalent could be drawn more than once.

Next, map the randomized positions onto the real sc.lengths for cell cell observation.

### Map normalized positions onto real SC

```{r map.rand2real, echo=FALSE}

#need sc length from the original
Position.DF$raw.foci.pos <- Position.DF$SC.length *Position.DF$value

male.whole.cell.mini$raw.rand_Foci1 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci1
male.whole.cell.mini$raw.rand_Foci2 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci2
male.whole.cell.mini$raw.rand_Foci3 <- male.whole.cell.mini$SC.length *male.whole.cell.mini$rand.Foci3

male.whole.cell.mini$rand.Foci2 <- as.numeric(male.whole.cell.mini$rand.Foci2 )
male.whole.cell.mini$rand.Foci3 <- as.numeric(male.whole.cell.mini$rand.Foci3)

#calq telo distnce
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

for(y in 1:length(male.whole.cell.mini)  ){
  male.whole.cell.mini$rbar[y] <- calq.intra.rbar(male.whole.cell.mini[y,])
}

for(rr in 1:length(male.whole.cell.mini)  ){
  male.whole.cell.mini$rand.rbar[rr] <- calq.intra.rbar_rand(male.whole.cell.mini[rr,])
}

mapp.cell.wide <- add_mouse(mapp.cell.wide)
mapp.cell.wide <- add_strain(mapp.cell.wide)

#cell wide telomere covariance bias
mapp.cell.wide <-  ddply(.data=male.whole.cell.mini,
                 .(fileName, hand.foci.count),
                 summarize,
                 nbiv = length(SC.length),
                 tot.SC = sum(SC.length),
                 mean.pos = mean(Foci1.PER),
                 mean.rand.pos = mean(rand.Foci1),
                 mean.raw = mean(c(Foci1,Foci2, Foci3), na.rm=TRUE ),
                 mean.raw.rand = mean( c(raw.rand_Foci1,raw.rand_Foci2,raw.rand_Foci3), na.rm=TRUE ),

                var.raw = var(c(Foci1,Foci2, Foci3), na.rm=TRUE ),
                 var.raw.rand = var( c(raw.rand_Foci1,raw.rand_Foci2,raw.rand_Foci3), na.rm=TRUE ),

                mean_rbar = mean(rbar),
                mean_rand.rbar = mean(rand.rbar)
)

maisel <- ggplot(mapp.cell.wide, aes(x=mean_rbar, y=mean_rand.rbar))+geom_point()+xlim(c(0,.2))+ylim(c(0,.2))

oopp <- ggplot(mapp.cell.wide, aes(y=mean.raw, x=mean.raw.rand, color=mouse) ) +geom_point()+ggtitle("average foci position by cell, real vs randomized")
oopp

hh <- ggplot(mapp.cell.wide, aes(y=var.raw,x=var.raw.rand) ) +geom_point()+ggtitle("variance in foci position (per cell), between real and randomized")

#what is the best way to plot these?

#gg <- ggplot(Position.DF, aes(y=raw.foci.pos, x=variable, color= as.factor(hand.foci.count) ))+geom_jitter()+ggtitle("raw foci positions, reals and rand")
#gg

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

distributions don't look different. 
Since I randomized the cell effect -- I need to map back into a cell wide summary to effectively compare the differences.

NEED to Calculate cell-wide telomere bias (use a ddplyr thing)

some of the plots compare the cell wide measures from real and randomized (I think this difference is what I should calculate for permuting or randomizing)

points closer to 1:1 line have foci position features near the average of the group
